---
title: "Genomic Ranges - GRanges"
author: "Sasha Ajay Malkani"
date: "2025-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing the IRanges package

```{r Installing the IRanges package}
# BiocManager::install("GenomicRanges")

```

## 2. load IRanges Package

```{r load IRanges Package}
library(GenomicRanges)

```

## 3. Example of GRanges

```{r Example of GRanges}
gr = GRanges(seqnames = c("chr1"), strand = c("+","-","+"), ranges = IRanges(start = c(1,3,5), width = 3))
gr

```

## 4. Flanking sequence on GRange is relative to the direction of transcription

```{r Flanking sequence on GRange is relative to the direction of transcription}
flank(gr, 5)

```

## 5. Promoters

```{r Promoters}
promoters(gr)

```

## 6. Sequence Information

```{r Sequence Information}
seqinfo(gr)

```

## 7. Give the Chromosome a length and then request for Sequence information

```{r Give the Chromosome a length and then request for Sequence information}
seqlengths(gr) = c("chr1" = 10)
seqinfo(gr)

```

## 8. Chromosome name given by seqlevels() function

```{r Chromosome name given by seqlevels() function}
seqlevels(gr)

```

## 9. Gaps() function covers everything not covered by IRanges and GRanges

```{r Gaps() function covers everything not covered by IRanges and GRanges}
gaps(gr)

```

## 10. Adding more Chromosomes

```{r Adding more Chromosomes}
# seqnames(gr) = c("chr1","chr2","chr1")
# Error Message invoked!
# Error in .normalize_seqnames_replacement_value(value, x) : 
  # levels of supplied 'seqnames' must be identical to 'seqlevels(x)'
```

## 11. Use Seqlevels() to create two levels then assign seq values to it

```{r Use Seqlevels() to create two levels then assign seq values to it}


```

## 12. Re-organize the seqlevels so "chr1" comes before "chr2" then sort

```{r Re-organize the seqlevels so "chr1" comes before "chr2" then sort}
  # present order of chromosomes
  seqlevels(gr)
  # re-organise chromosomes so chr2 comes before chr1
  seqlevels(gr) = c("chr2","chr1")
  # Then sort order of chromosomes
  sort(gr)

```

## 13. Assigning a genome name

```{r Assigning a genome name}
genome(gr) = "hg19"
gr

```

## 14. Displaying genome information of Seqinfo()

```{r Displaying genome information of Seqinfo()}
seqinfo(gr)

```

## 15. Make a copy of the genome

```{r Make a copy of the genome}
gr2 = gr
genome(gr2) = "hg18"

```

## 16. Find Overlaps between genome "hg19" and "hg18"

```{r Find Overlaps between genome "hg19" and "hg18"}
# findOverlaps(gr, gr2)
# Error Message invoked!
# Error in mergeNamedAtomicVectors(genome(x), genome(y), what = c("sequence",  : 
  # sequences chr2, chr1 have incompatible genomes:
  # - in 'x': hg19, hg19
  # - in 'y': hg18, hg18

```






