---
title: "Genomic Ranges - Basic GRanges Usage"
author: "Sasha Ajay Malkani"
date: "2025-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing the IRanges package

```{r Installing the IRanges package}
# BiocManager::install("GenomicRanges")

```

## 2. Load GenomicRanges

```{r Load GenomicRanges}
library(GenomicRanges)

```

## 3. Construct IRanges and a data frame

```{r Construct IRanges and a data frame}
ir = IRanges(start = 1:3, width = 2)
df = DataFrame(ir = ir, score = rnorm(3))
df

```

## 4. Subset IRange of length = 1

```{r Subset IRange}
df[1,1]

```

## 5. Use the $ operator to print data frame

```{r Use the $ operator to print data frame}
df$ir

```

## 6. Classic Data Frame in R with IRanges

```{r Classic Data Frame in R with IRanges}
df2 = data.frame(ir = ir)
df2

```

## 7. Constructing GRanges

```{r Constructing GRanges}
gr <- GRanges(seqnames = "chr1", strand = c("+","-","+"), ranges = IRanges(start = c(1,3,5), width = 3))
gr

```

## 8. GRanges have something called values

```{r GRanges have something called values}
values(gr) = DataFrame(score = rnorm(3))
gr

```

## 9. Accessing the values column in GRanges (gr)

```{r Accessing the values column in GRanges (gr)}
values(gr)

```

## 10. Accessing the values column with mcols

```{r Accessing the values column with mcols}
mcols(gr)

```

## 11. Access a different column using dollar operator

```{r Access a different column using dollar operator}
gr$score

```

## 12. Assigning a new column

```{r Assigning a new column}
gr$score2 = gr$score / 3
gr

```

## 13. New GRanges

```{r New GRanges}
gr2 <- GRanges(seqnames = c("chr1", "chr2", "chr1"), strand = "*", ranges = IRanges(start = c(1,3,5), width = 3))
gr2

```

## 14. Old GRanges

```{r Old GRanges}
gr

```

## 15. Find Overlaps between gr and gr2

```{r Find Overlaps between gr and gr2}
findOverlaps(gr, gr2)

```

## 16. Ignoring Strand in findOverlaps() function

```{r Ignoring Strand in findOverlaps() function}
findOverlaps(gr, gr2, ignore.strand = TRUE)

```

## 17. Subset Overlaps

```{r Subset Overlaps}
subsetByOverlaps(gr, gr2)

```

## 18. Subset Overlaps (reverse)

```{r Subset Overlaps (reverse)}
subsetByOverlaps(gr2, gr)

```

## 19. Construct classic R data frame

```{r Construct classic R data frame}
df = data.frame(chr = "chr1", start = 1:3, end = 4:6, score = rnorm(3))
df

```

## 20. Converting classic R data frame to a GRanges

```{r Converting classic R data frame to a GRanges}
makeGRangesFromDataFrame(df)

```

## 21. Converting classic R data frame to GRanges keeping extra columns

```{r Converting classic R data frame to GRanges keeping extra columns}
makeGRangesFromDataFrame(df, keep.extra.columns = TRUE)

```





