---
title: "R Base Types"
author: "Sasha Ajay Malkani"
date: "2025-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Atomic Vectors (same data type) or Vectors

```{r Atomic Vectors (same data type) or Vectors}
x = 1:10
x

```

## 2. Vectors can have names

```{r Vectors can have names}
names(x) = letters[1:10]
class(x)

```

## 3. Subset Vectors

```{r Subset Vectors}
x[1:3]

```

## 4. Index based on its names

```{r Index based on its names}
x[c("a","b")]

```

## 5. Names do not have to be unique on vectors

```{r Names do not have to be unique on vectors}
x = 1:3
names(x) = c("a","a","b")
x
x["a"]

```

## 6. Check for duplicates

```{r Check for duplicates}
anyDuplicated(names(x))

```

## 7. Second element in the names vector is a duplicate

```{r Second element in the names vector is a duplicate}
names(x)

```

## 8. No duplicates

```{r No duplicates}
names(x) = letters[1:3]
names(x)
anyDuplicated(names(x))

```

## 9. The difference between Numerics and Integers

```{r The difference between Numerics and Integers}
x = 1
class(x)

x = 1:3
class(x)

```

## 10. Getting a single integer with command line

```{r Getting a single integer with command line}
x = 1L
class(x)

```

## 11. Machine limit to big integers

```{r Machine limit to big integers}
.Machine$integer.max
2^31 - 1 == .Machine$integer.max

```

## 12. A big number BUT slightly smaller than the human genome!

```{r A big number BUT slightly smaller than the human genome!}
round(.Machine$integer.max / 10^6)

```

## 13. Convert from integer to numeric data type to avoid the problem of integer limits

```{r 13. Convert from integer to numeric data type}
as.numeric(1L)

```

## 14. Matrices - Constructing a matrix

```{r Matrices - Constructing a matrix}
x = matrix(1:9, ncol = 3, nrow = 3)
x

```

## 15. Rownames

```{r Rownames}
rownames(x) = letters[1:3]
x

```

## 16. Dimensions of a matrix

```{r Dimensions of a matrix}
dim(x)

```

## 17. Number of rows and columns in a matrix

```{r Number of rows and columns in a matrix}
nrow(x)
ncol(x)

```

## 18. First two rows

```{r First two rows}
x[1:2,]

```

## 19. First two columns

```{r First two columns}
x[,1:2]

```

## 20. Both at the same time

```{r Both at the same time}
x[1:2,1:2]

```

## 21. Sub-setting using characters

```{r Sub-setting using characters}
x["a",]

```

## 22. One dimensional Matrix

```{r One dimensional Matrix}
x["a",, drop=FALSE]

```

## 23. Sub-setting Matrices

```{r Sub-setting Matrices}
x[x>5]

```

## 24. Matrices - Constructing a matrix - fills up columns first

```{r Matrices - Constructing a matrix - fills up columns first}
x = matrix(1:9, ncol = 3, nrow = 3)
x

```

## 25. Constructing a matrix filling by row first

```{r Constructing a matrix filling by row first}
x = matrix(1:9, ncol = 3, nrow = 3, byrow = TRUE)
x

```

## 26. Constructing a list of different data types

```{r Constructing a list of different data types}
x = list(a = rnorm(3), b=letters[1:5], matrix)
x

```

## 27. Sub-setting the first two elements in a list

```{r Sub-setting the first two elements in a list}
x[1:2]

```

## 28. Sub-setting the first element

```{r Sub-setting the first element}
x[1]

```

## 29. Extracting the first element itself (referencing 2D array)

```{r 29. Extracting the first element itself}
x[[1]]

```

## 30. Subsetting the first element by name (classic)

```{r Subsetting the first element by name (classic)}
x["a"]

```

## 31. Subsetting the first element by name (using dollar operator)

```{r Subsetting the first element by name (using $ operator)}
x$a

```

## 32. Partial Matching

```{r Partial Matching}
names(x) = c("a", "letters", "c")
x$letters
x$let

```

## 33. Partial Matching can be dangerous

```{r Partial Matching can be dangerous}
names(x) = c("a", "letters", "letters2")
x$let

```

## 34. Single bracket notation avoids partial matching and potential errors

```{r Single bracket notation}
names(x) = c("a", "letters", "letters2")
x["letters"]

```

## 35. Constructing a list where each element in the list is a single number

```{r Constructing a list where each element in the list is a single number}
as.list(1:3)

```

## 36. Construct a list where each element of the list is of the same type

```{r Construct a list where each element of the list is of the same type}
x = list(rnorm(3), 3:9)
x

```

## 37. Using lapply() function

```{r Using lapply() function}
lapply(x, mean)

```

## 38. Generating a vector instead of a list

```{r Generating a vector instead of a list}
unlist(lapply(x, mean))

```

## 39. simplifying lapply() output with sapply() function

```{r simplifying lapply() output with sapply() function}
sapply(x, mean)

```

## 40. Constructing a dataframe

```{r Constructing a dataframe}
x <- data.frame(sex = c("M", "M", "F"), age = c(32, 34, 29))
x

```

## 41. Selecting columns in a data frame

```{r Selecting columns in a data frame}
x$sex

```

## 42. Using single or double brackets to select a specific column

```{r Using single or double brackets to select a specific column}
x[["sex"]]

```

## 43. Sub-setting the data frame the same way we do for matrices

```{r Sub-setting the data frame the same way we do for matrices}
x[1, "sex"]

```

## 44. Row names in a data frame have to be unique (1,2,3)

```{r Row names in a data frame have to be unique}
x

```

## 45. Using sapply and lapply on dataframes

```{r Using sapply and lapply on dataframes}
sapply(x, class)

```

## 46. Converting and data frame into a matrix (character matrix)

```{r Converting and data frame into a matrix (character matrix)}
as.matrix(x)

```

## 47. Converting a matrix into a list

```{r Converting a matrix into a list}
as.list(x)

```

## 48. Conversions in Bioconductor use the as() function from the methods package

```{r 48. Conversions in Bioconductor from the methods package}
library(methods)
as(x, "matrix")

```


