---
title: "BioMART"
author: "Sasha Ajay Malkani"
date: "2025-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing Bioconductor package

```{r 1. Installing Bioconductor package}
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install()

```

## 2. Installing "biomaRt"

```{r 2. Installing "biomaRt"}
BiocManager::install(c("biomaRt"), force = TRUE)

```

## 3. Troubleshooting installing packages

```{r 3. Troubleshooting installing packages}
BiocManager::valid()

```

## 4. Load Libraries

```{r 4. Load Libraries}
library(GenomicRanges)
library(biomaRt)
library(Biobase)
library(GenomicFiles)

```

## 5. Listing a marts

```{r 5. Listing a marts}
head(listMarts())

```

## 6. Picking Ensembl database

```{r 6. Picking Ensembl database}
mart <- useMart("ensembl")
mart

```

## 7. List datasets in the database

```{r 7. List datasets in the database}
head(listDatasets(mart))

```

## 8. Selecting a dataset

```{r 8. Selecting a dataset}
ensembl <- useDataset("hsapiens_gene_ensembl", mart)
ensembl

```

## 9. Building a query

```{r 9. Building a query}
values <- c("202763_at","209310_s_at","207500_at")
getBM(attributes = c("ensembl_gene_id", "affy_hg_u133_plus_2"),
                     filters = "affy_hg_u133_plus_2", values = values, mart = ensembl)

```

## 10. Calling list attributes on Ensembl database

```{r 10. Calling list attributes on Ensembl database}
attributes <- listAttributes(ensembl)
nrow(attributes)

```

## 11. Getting names of the attributes

```{r 11. Getting names of the attributes}
head(attributes)

```

## 12. Getting names of the attributes (bottom)

```{r 12. Getting names of the attributes (bottom)}
tail(attributes)

```

## 13. Getting names of the attributes (last 100) output not shown

```{r 13. Getting names of the attributes (last 100) output not shown}
tail(attributes, n=100)

```

## 14. Getting names of the attributes (last 500) output not shown

```{r 14. Getting names of the attributes (last 500) output not shown}
tail(attributes, n=500)

```

## 15. Listing filters

```{r 15. Listing filters}
filters <- listFilters(ensembl)
head(filters)

```

## 16. Number of filters

```{r 16. Number of filters}
nrow(filters)

```

## 17. Attribute pages

```{r 17. Attribute pages}
attributePages(ensembl)

```

## 18. Listing Attributes without homologs

```{r 18. Listing Attributes without homologs}
attributes <- listAttributes(ensembl, page = "feature_page")
head(attributes)

```

