---
title: "Expression Set"
author: "Sasha Ajay Malkani"
date: "2025-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. install ALL package

```{r 1. install ALL package}
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")

BiocManager::install("ALL")

```

## 2. Load ALL package

```{r 2. Load ALL package}
library(ALL)

```

## 3. Print, display ALL package

```{r 3. Print, display ALL package}
data(ALL)
ALL

```

## 4. Details of Experiment given using an accessor function

```{r 4. Details of Experiment given using an accessor function}
experimentData(ALL)

```

## 5. More information can be obtained looking at the help page

```{r 5. More information can be obtained looking at the help page}
?ALL

```

## 6. Accessing the expression matrix

```{r 6. Accessing the expression matrix}
exprs(ALL)[1:4, 1:4]

```

## 7. Details of Sample names by the sample names accessor function

```{r 7. Details of Sample names by the sample names accessor function}
head(sampleNames(ALL))

```

## 8.  Details of Feature names by the feature names accessor function

```{r 8.  Details of Feature names by the feature names accessor function}
head(featureNames(ALL))

```

## 9. Accessing phenotype data from the data frame

```{r 9. Accessing phenotype data from the data frame}
head(pData(ALL))

```

## 10. Accessing specific covariants for modelling purposes

```{r 10. Accessing specific covariants for modelling purposes}
pData(ALL)$sex

```

## 11. Alternative

```{r 11. Alternative}
ALL$sex

```

## 12. Selecting the first 5 samples

```{r 12. Selecting the first 5 samples}
ALL[, 1:5]

```

## 13. Selecting the first 10 features

```{r 13. Selecting the first 10 features}
ALL[1:10,]

```

## 14. Selecting both features and samples

```{r 14. Selecting both features and samples}
ALL[1:10, 1:5]

```

## 15. Obtaining Feature data (normally not stored in the object itself)

```{r 15. Obtaining Feature data (normally not stored in the object itself)}
featureData(ALL)

```

## 16. Getting the first 5 feature names

```{r 16. Getting the first 5 feature names}
ids = featureNames(ALL)[1:5]
ids

```

## 17. Mapping the Affymetrix ID's into something specific

### 17. A. Install hgu95av2.db package

```{r 17. A. Install hgu95av2.db package}
BiocManager::install("hgu95av2.db")

```

### 17. B. Loading hgu95av2.db package

```{r 17. B. Loading hgu95av2.db package}
library(hgu95av2.db)

```

### 17. C. Mapping Affymetrix ID's into ENTREZID

```{r 17. C. Mapping Affymetrix IDs into ENTREZID}
as.list(hgu95av2ENTREZID[ids])

```

## 18. Using PhenoData instead of pData (the recommended one!)

```{r 18. Using PhenoData instead of pData (the recommended one!)}
phenoData(ALL)

```

## 19. Better example

```{r 19. Better example}
names(pData(ALL))

```

## 20. VarLabels

```{r 20. VarLabels}
varLabels(ALL)

```

## 21. PhenoData is different to pData

```{r 21. PhenoData is different to pData}
phenoData(ALL)

# vs
head(pData(ALL))

```

## 22. PhenoData contains a pData slot

```{r 22. PhenoData contains a pData slot}
head(pData(phenoData(ALL)))

```



