---
title: "R S4 Classes"
author: "Sasha Ajay Malkani"
date: "2025-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing Bioconductor package

```{r 1. Installing Bioconductor package}
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install()

```

## 2. Installing "ALL" and "Genomic Ranges Packages"

```{r 2. Installing "ALL" and "Genomic Ranges Packages"}
BiocManager::install(c("ALL", "GenomicRanges"), force = TRUE)

```

## 3. Load the Libraries

```{r 3. Load the Libraries}
library(ALL)
library(GenomicRanges)

```

## 4. LM Object from a linear model

```{r 4. LM Object from a linear model}
df <- data.frame(y = rnorm(10), x = rnorm(10))
lm.object <- lm(y ~ x, data = df)
lm.object

```

## 5. Asking for the objects class

```{r 5. Asking for the objects class}
class(lm.object)

```

## 6. Using names() function on the S4 object

```{r 6. Using names() function on the S4 object}
names(lm.object)

```

## 7. Create a list of two elements letters and numbers

```{r 7. Create a list of two elements letters and numbers}
xx = list(a = letters[1:3], b = rnorm(4))
xx

```

## 8. Class XX = lm

```{r 8. Class XX equals lm}
class(xx) = "lm"

```

## 9. Print the lm object

```{r 9. Print the lm object}
xx

```

## 10.Load library

```{r 10.Load library}
library(ALL)

```

## 11. Load dataset

```{r 11. Load dataset}
data(ALL)

```

## 12. Print dataset

```{r 12. Print dataset}
ALL

```

## 13. Calling Class of the object

```{r 13. Calling Class of the object}
class(ALL)

```

## 14. Helper function isS4() determines whether an object is of the S4 class

```{r 14. Helper function isS4() determines whether an object is of the S4 class}
isS4(ALL)

```

## 15. Getting help with an S4 class object

```{r 15. Getting help with an S4 class object}
class?ExpressionSet

```

## 16. Alternative for help with S4 class object

```{r 16. Alternative for help with S4 class object}
?"ExpressionSet-class"

```

## 17. Use constructors all the time in R e.g. list() function

```{r 17. Use constructors all the time in R e.g. list() function}
xx = list(a = 1:3)

```

## 18. Constructor for the expression set

```{r 18. Constructor for the expression set}
ExpressionSet()

```

## 19. Calling for help with an S4 object gives information on the object and its constructor

```{r 19. Calling for help with an S4 object gives information on the object and its constructor}
?ExpressionSet

```

## 20. Classic way of defining an Expression set

```{r 20. Classic way of defining an Expression set}
new("ExpressionSet")

```

## 21. Defining a S4 class

```{r 21. Defining a S4 class}
getClass("ExpressionSet")

```

## 22. Using the slots with the @ symbol to access data in the expression set

```{r 22. Using the slots with the @ symbol to access data in the expression set}
ALL@annotation

```

## 23. Adjusting screen options

```{r 23. Adjusting screen options}
opar = options(width = 80)

```

## 24. Class of expression set

```{r 24. Class of expression set}
getClass("ExpressionSet")

```

## 25. Using the slots with the @ symbol to access data in the expression set

```{r 25. Using the slots with the @ symbol to access data in the expression set}
ALL@annotation

```

## 26. Alternative way of accessing data in a S4 object expression set

```{r 26. Alternative way of accessing data in a S4 object expression set}
slot(ALL, "annotation")

```

## 27. Alternative way of accessing data in a S4 object expression set using an accessor function

```{r 27. Alternative way of accessing data in a S4 object expression set using an accessor function}
annotation(ALL)

```

## 28. Accessing the help for the class

```{r 28. Accessing the help for the class}
?ExpressionSet

```

## 29. Updating old objects with newer versions

```{r 29. Updating old objects with newer versions}
# OLD_OBJECT = updateObject(OLD_OBJECT)

```

## 30. Testing that an object is valid

```{r 30. Testing that an object is valid}
validObject(ALL)

```
