---
title: "Limma"
author: "Sasha Ajay Malkani"
date: "2025-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing Bioconductor package

```{r 1. Installing Bioconductor package}
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install()

```

## 2. Installing "limma, leukemiasEset"

```{r 2. Installing "limma, leukemiasEset"}
BiocManager::install(c("limma", "leukemiasEset"), force = TRUE)

```

## 3. Troubleshooting installing packages

```{r 3. Troubleshooting installing packages}
BiocManager::valid()

BiocManager::install(c(
  "BiocParallel", "oligo", "promises", "reticulate", "Rsamtools"
), update = TRUE, ask = FALSE, force = TRUE)

```

## 4. Load Libraries

```{r 4. Load Libraries}
library(limma)
library(leukemiasEset)

```

## 5. Displaying the leukemiasEset expression set

```{r 5. Displaying the leukemiasEset expression set}
data(leukemiasEset)
leukemiasEset

```

## 6. Profiling the subsets of Leukemia

```{r 6. Profiling the subsets of Leukemia}
table(leukemiasEset$LeukemiaType)

```

## 7. Subset the Leukemia types (select ALL & NoL)

```{r 7. Subset the Leukemia types (select ALL & NoL)}
ourData <- leukemiasEset[, leukemiasEset$LeukemiaType %in% c("ALL", "NoL")]

```

## 8. Leukemia type is a factor

```{r 8. Leukemia type is a factor}
ourData$LeukemiaType

```

## 9. Getting rid of the three redundant factors in leukemia type not in the subset

```{r 9. Getting rid of the three redundant factors in leukemia type not in the subset}
ourData$LeukemiaType <- factor(ourData$LeukemiaType)
ourData$LeukemiaType

```

## 10. Setting up the design matrix

```{r 10. Setting up the design matrix}
design <- model.matrix(~ ourData$LeukemiaType)
head(design)

```

## 11. Fitting a linear model using (limma) to all the genes separately

```{r 11. Fitting a linear model using (limma) to all the genes separately}
fit <- lmFit(ourData, design)

```

## 12. Fit the model

```{r 12. Fit the model}
fit <- eBayes(fit)

```

## 13. Look at the top of the table

```{r 13. Look at the top of the table}
topTable(fit)

```

## 14. Getting rid of the three redundant factors in leukemia type not in the subset

```{r 14. Getting rid of the three redundant factors in leukemia type not in the subset}
ourData$LeukemiaType <- factor(ourData$LeukemiaType)
ourData$LeukemiaType

```

## 15. Get the first gene from the table

```{r 15. Get the first gene from the table}
topTable(fit, n = 1)

```

## 16. Get the gene name

```{r 16. Get the gene name}
genename <- rownames(topTable(fit, n=1))
genename

```

## 17. Compute the mean expression inside each group

```{r 17. Compute the mean expression inside each group}
typeMean <- tapply(exprs(ourData)[genename,], ourData$LeukemiaType, mean)
typeMean

```

## 18. The log fold change between ALL and NoL

```{r 18. The log fold change between ALL and NoL}
typeMean["NoL"] - typeMean["ALL"]

```

## 19. Create design matrix 2

```{r 19. Create design matrix 2}
design2 <- model.matrix(~ ourData$LeukemiaType - 1)
head(design2)

```

## 20. Changing the names of the columns to shorter names

```{r 20. Changing the names of the columns to shorter names}
colnames(design2) <- c("ALL", "NoL")
head(design2)


```

## 21. Fit the model (as before)

```{r 21. Fit the model (as before)}
fit2 <- lmFit(ourData, design2)

```

## 22. Make contrasts

```{r 22. Make contrasts}
contrast.matrix <- makeContrasts("ALL-NoL", levels = design2)
contrast.matrix

```

## 23. Execute contrasts.fit

```{r 23. Execute contrasts.fit}
fit2C <- contrasts.fit(fit2, contrast.matrix)

```

## 24. Use eBayes command to borrow information across genes

```{r 24. Use eBayes command to borrow information across genes}
fit2C <- eBayes(fit2C)

```

## 25. Display results

```{r 25. Display results}
topTable(fit2C)

```

