---
title: "Short Read"
author: "Sasha Ajay Malkani"
date: "2025-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing Bioconductor package

```{r 1. Installing Bioconductor package}
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install()

```

## 2. Installing "ShortRead"

```{r 2. Installing "ShortRead"}
BiocManager::install(c("ShortRead"), force = TRUE)

```

## 3. Troubleshooting installing packages

```{r 3. Troubleshooting installing packages}
BiocManager::valid()

BiocManager::install(c(
  "BiocParallel", "promises", "Rsamtools"
), update = TRUE, ask = FALSE, force = TRUE)

```

## 4. Load the "ShortRead" package

```{r 4. Load the "ShortRead" package}
library(ShortRead)

```

## 5. Pointer down to a FASTQ file

```{r 5. Pointer down to a FASTQ file}
fastqDir <- system.file("extdata", "E-MTAB-1147", package = "ShortRead")
fastqPath <- list.files(fastqDir, pattern = ".fastq.gz$", full = TRUE)[1]
reads <- readFastq(fastqPath)
reads

```

## 6. Alternative instantiate a FASTQ file object

```{r 6. Alternative instantiate a FASTQ file object}
fqFile <- FastqFile(fastqPath)
fqFile

```

## 7. Reading the FASTQ file

```{r 7. Reading the FASTQ file}
reads <- readFastq(fqFile)

```

## 8. Accessing individual reads using sread

```{r 8. Accessing individual reads using sread}
sread(reads)[1:2]

```

## 9. Accessing Quality values for each read

```{r 9. Accessing Quality values for each read}
quality(reads)[1:2]

```

## 10. Accessing read names

```{r 10. Accessing read names}
id(reads)[1:2]

```

## 11. Converting (coercing) standard quality score into integers from 1 - 40

```{r 11. Converting (coercing) standard quality score into integers from 1 - 40}
as(quality(reads), "matrix")[1:2,1:10]

```







