---
title: "Biostrings"
author: "Sasha Ajay Malkani"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Installing Biostrings Package

```{r Installing Biostrings Package}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::valid()

BiocManager::install(c("BiocParallel", "Rsamtools"), update = TRUE, ask = FALSE, force = TRUE)

BiocManager::install("Biostrings")

```

## 2. Loading Biostrings Package

```{r Loading Biostrings Package}
library(BiocGenerics)
library(S4Vectors)
library(IRanges)
library(GenomicRanges)
library(Biostrings)

```

## 3. Representing sequences

```{r Representing sequences}
dna1 <- DNAString("ACGT-G")
dna1

```

## 4. DNA string set , a collection of DNA strings

```{r DNA string set , a collection of DNA strings}
dna2 <- DNAStringSet(c("ACGT", "GTCA", "GCTA"))
dna2

```

## 5. IUPAC code

```{r IUPAC code}
IUPAC_CODE_MAP

```

## 6. Subset DNA gives a subsequence

```{r }
dna1[2:4]

```

## 7. Subset a DNA String set -> another string set

```{r Subset a DNA String set -> another string set}
dna2[1:2]

```

## 8. Get a string out of a DNA string set

```{r Get a string out of a DNA string set}
dna2[[1]]

```

## 9. Put name on a DNA string set

```{r Put name on a DNA string set}
names(dna2) = paste0("seq", 1:3)
dna2

```

## 10. Number of bases in the string

```{r Number of bases in the string}
width(dna2)

```

## 11. Sort DNA string set

```{r Sort DNA string set}
sort(dna2)

```

## 12. Reversing a DNA string set

```{r Reversing a DNA string set}
rev(dna2)

```

## 13. Reversing a DNA string

```{r Reversing a DNA string}
rev(dna1)

```

## 14. Reversion of a DNA string set 

using the reverse() function, reverses each string in the DNA string set

```{r Reversion of a DNA string set}
reverse(dna2)

```

## 15. Use reverse compliment

```{r Use reverse compliment}
reverseComplement(dna2)

```

## 16. Translate DNA to protein

```{r Translate DNA to protein}
translate(dna2)

```

## 17. Alphabet frequency gives a general frequency table of DNA bases

```{r Alphabet frequency gives a general frequency table of DNA bases}
alphabetFrequency(dna2)

```

## 18. Counting GC content of a string

```{r Counting GC content of a string}
letterFrequency(dna2, letters = "GC")

```

## 19. Higher Order Nucleotide frequency e.g. dinucleotides

```{r Higher Order Nucleotide frequency e.g. dinucleotides}
dinucleotideFrequency(dna2)

```

## 20. Consensus Matrix

```{r Consensus Matrix}
consensusMatrix(dna2)

```

