---
title: "Genomic Ranges - Lists"
author: "Sasha Ajay Malkani"
date: "2025-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load library

```{r Load library}
library(GenomicRanges)

```

## 2. Constructing a GRanges list

```{r Constructing a GRanges list}
gr1 <- GRanges(seqnames = "chr1", ranges = IRanges(start = 1:4, width = 3))
gr2 <- GRanges(seqnames = "chr2", ranges = IRanges(start = 1:4, width = 3))
gL <- GRangesList(gr1 = gr1, gr2 = gr2)
gL

```

## 3. Indexing the 1st element in the GRanges list using [[]]

```{r Indexing the 1st element in the GRanges list using [[]]}
gL[[1]]

```

## 4. Indexing the 1st element in the GRanges list using the $ operator

```{r Indexing the 1st element in the GRanges list using the $ operator}
gL$gr1

```

## 5. Using the start() access function

```{r Using the start() access function}
start(gL)

```

## 7. How long are each of the elements in GRanges (function out of date)

```{r How long are each of the elements in GRanges (function out of date)}
# First do this
elementLengths=elementNROWS

# Then run the function
elementLengths(gL)

```

## 8. Using sapply() function to find the length of each element in GRanges (slower)

```{r Using sapply() function to find the length of each element in GRanges (slower)}
sapply(gL, length)

```

## 9. Indoor Applies

```{r Indoor Applies}
shift(gL, 10)

```

## 10. FindOverlaps() function between GRanges List (gL) and Grange (gr2)

```{r 10. FindOverlaps() function between GRanges List (gL) and Grange (gr2)}
findOverlaps(gL, gr2)

```

## 11. FindOverlaps() function between GRanges List (gL) and GRanges List (gL)

```{r 11. FindOverlaps() function between GRanges List (gL) and GRanges List (gL)}
findOverlaps(gL, gL)

```

